---
title  : 'Programming assignment 4'
output: github_document

---

**Author**: Yhosep Barba  
**Date**: Last update: `r Sys.time()`


# Overview

<!-- 
  Talk briefly about what you did here 
  Describe your hypotheses
-->

Monolingual English speakers would have a higher VOT.
Bilinguals would have a shorter or similar VOT to monolingual Spanish Speakers (if existed). They would also show more control over F1 and F2 frequencies, exhibiting a broader vowel space in their pronounciation skills. All of this taking into account previous and more exposure to the Spanish language. 


# Prep

## Libraries 

```{r}
library(tidyverse)

```

## Load data

```{r}
df <- list.files(path="../data", full.names= TRUE) %>%
  lapply(read_csv) %>%
  bind_rows() 

```

## Tidy data
```{r}
df2 <- df %>%
  separate("fileID", c("ID","Word")) %>%
  select(-c(notes,data)) %>%
  na.omit()
```
# Analysis

## Descriptives

```{r}
means <- df2 %>%
  group_by(ID) %>%
  summarize(avg_f1 = mean(f1),
            sd_f1 = sd(f1),
            avg_f2 = mean(f2),
            sd_f2 = sd(f2),
            avg_vot = mean(vot),
            sd_vot = sd(vot))
 knitr::kable(means) 
```

## Visualization

```{r}
df2 %>%
   ggplot() +
   aes(x = ID, y = f1) +
   geom_boxplot() +
   labs(x = "ID", y = "f1") 

df2 %>%
   ggplot() +
   aes(x = ID, y = f2) +
   geom_boxplot() +
   labs(x = "ID", y = "f2")

df2 %>%
   ggplot() +
   aes(x = ID, y = vot) +
   geom_boxplot() +
   labs(x = "ID", y = "vot")
```

<!-- 
Also include a professional looking figure illustrating an example of the acoustics 
of the production data, i.e., a plot generated in praat.
You decide what is relevant (something related to your hypothesis). 
Think about where this file should be located in your project. 
What location makes most sense in terms of organization? 
How will you access the file (path) from this .Rmd file?
If you need help consider the following sources: 
  - Search 'Rmarkdown image' on google, stackoverflow, etc.
  - Search the 'knitr' package help files in RStudio
  - Search the internet for HTML code (not recommended, but it works)
  - Check the code from my class presentations (may or may not be helpful)
-->
![An image](/Users/diegomedina/Desktop/pa_4/soundwave.pdf)


## Hypothesis test

```{r}
#| label: stats
# Conduct a simple statistical analysis here (optional)

```


# Conclusion

<!-- 
Revisit your hypotheses (refer to plots, figures, tables, statistical tests, 
etc.)

Reflect on the entire process. 
What did you enjoy? What did you hate? What did you learn? 
What would you do differently?
-->

To conclude, there is variability in the average F1 and F2 frequencies among the six participants. L2 learners and bilinguals exhibit different patterns of vowel production. I believe that with only six participants, it is difficult to determine and/or categorize vowel formants. On the other hand, VOT results matched with the hypothesis: bilinguals portray a shorter VOT compared to L2 learners. Further information would be required if we want to include exposure as an independent variable. 

I enjoyed the process, I believe it was really interesting to see how, through some scripts, you are able to segment different audios (this saves a lot of time). I still struggle a little bit in understanding how codes and scripts work. I guess I would have to instruct myself during summer and find the logic behind it. In addition to this, I find R very useful, although I hate that even small mistakes do not allow to run some instructions. If I had to do something different, I would include colors to my graphs. 

</br></br>
